Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States

### 2024-11-04 paper/paper.qmd
remove taboo words

---
title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: >
  This study employs a polls-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college. By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris. Our analysis also reveals a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore the value of aggregating polls over relying on individual surveys, offering a more comprehensive and robust forecast of electoral outcomes.
format: 
  pdf:
    number-sections: true
    header-includes:
      - \usepackage{pdfpages}
      - '\KOMAoption{captions}{tableheading}'
      - \usepackage[margin=1in]{geometry}

@@

This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models. By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact. This could provides insights for political strategists, media analysts, and the general public by offering a nuanced view of the electoral landscape.

@@

## Explorations of Variables of Interest

@@

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-ster
#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls for balanced insights.'

@@

@fig-pie illustrates the distribution of polls between candidates and poll types. The left chart shows that the majority of polls are conducted at the state level, with a smaller portion being national. The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump. This distribution underscores the model’s balanced approach to capturing state-level nuances as well as broader national trends, providing a comprehensive view of candidate support across different contexts

@@

# Filter out Alaska, Hawaii, and Puerto Rico using subset
map_data_continental <- subset(map_data, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

ggplot(map_data_continental) +
  geom_sf(aes(fill = kamala_proportion * 100), color = "white", size = 0.2) +
  geom_text(aes(label = STUSPS, geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust color and size as needed
  scale_fill_distiller(palette = "RdBu", direction = 1, na.value = "grey", 
                       name = "Harris's Mean Support in Polls (%)") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Make the plot wider
  )

@@

In our approach, we will employ linear modeling of voter support percentage (pct) on pollster and other independent variables such as sample size, poll recency, and poll scope (state vs. national). This will allow us to smooth out the inherent noise, biases, and variability across different pollsters. Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome. Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election. This approach helps us capture both national trends and critical state-level dynamics.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-causal-model
#| fig-cap: "Factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."
# Load and display the saved PNG
knitr::include_graphics(here("other", "graphs", "polls_of_polls_dag.png"))

@@

In our model, we aim to smooth out discrepancies and biases across various polling organizations using a polls-of-polls approach. Given the potential for individual pollsters to introduce systematic differences—due to variations in sampling methods, question phrasing, and historical leanings—our model includes a pollster variable to adjust for these organization-specific biases. This allows us to capture an aggregated view of public support that is less susceptible to the idiosyncrasies of any single poll. Furthermore, we incorporate sample size as a predictor, as polls with larger samples tend to yield more stable and reliable results, reducing random fluctuations caused by smaller samples. The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is crucial for understanding the nuanced political landscape. Additionally, recency is included to prioritize more recent polls, as public opinion can shift rapidly in response to political events, and recent data is generally more reflective of current sentiment. By integrating these factors, our model seeks to produce a more stable and comprehensive measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary
#| tbl-cap: "Model performance summary showing improved fit and accuracy as State and Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in Model 3, and RMSE decreasing from 3.053 to 1.836." 
# Load your models
model1 <- readRDS(here("models/model1.rds"))
model2 <- readRDS(here("models/model2.rds"))
model3 <- readRDS(here("models/model3.rds"))

# Extract summary statistics and add variables included in each model
model_summary <- tibble::tibble(
  Model = c("Model 1", "Model 2", "Model 3"),
  Variables = c("Pollster, Sample Size", "Pollster, Sample Size, State", "Pollster, Sample Size, State, Recency"),
  `R²` = c(summary(model1)$r.squared, summary(model2)$r.squared, summary(model3)$r.squared),
  `Adjusted R²` = c(summary(model1)$adj.r.squared, summary(model2)$adj.r.squared, summary(model3)$adj.r.squared),
  `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
  `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
  `RMSE` = c(sqrt(mean(residuals(model1)^2)), sqrt(mean(residuals(model2)^2)), sqrt(mean(residuals(model3)^2)))
)

@@

Aggregating support across all polls and applying our weighted approach, we estimate the overall support for each candidate. The weights, calculated based on both the poll's numeric grade and poll score, ensure that higher-quality polls with less bias contribute more to our estimates. Based on this approach, Kamala Harris has an estimated overall weighted support of around 47.77%, while Donald Trump stands at around 46.31%. This aggregation provides a comprehensive view, taking into account the varied methodologies and sampling qualities of different polling organizations.

@@

# Step 1: Subset data for Kamala Harris, calculate weights, join with fitted values from model3, and calculate weighted_pct
kamala_data <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
    fitted_values = fitted(model3),  # Replace this with actual fitted values for Kamala
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Kamala
  )

@@

# Step 2: Subset data for Donald Trump, calculate weights, join with fitted values from model6, and calculate weighted_pct
trump_data <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
             sum(numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore), na.rm = TRUE),
    fitted_values = fitted(model6),  # Replace this with actual fitted values for Trump
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Trump
  )

# Calculate the overall weighted prediction for Donald Trump
trump_weighted_support <- sum(trump_data$weighted_pct, na.rm = TRUE)

@@

For each state, we compare the predicted support percentages and determine the projected winner based on which candidate has higher weighted support. Additionally, we calculate the margin of difference between the candidates, which can provide insights into how close each state’s race is.

@@

# Step 1: Subset data for Kamala Harris, join fitted values, then calculate weights within each state
kamala_data_a2 <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(fitted_values = fitted(model3)) %>%  # Add fitted values from model 3 for Kamala Harris
  group_by(state) %>%
  mutate(
    weight = numeric_grade * (max_pollscore - pollscore) /
             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
  ) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_kamala = sum(fitted_values * weight, na.rm = TRUE))

# Step 2: Repeat for Donald Trump with model6
trump_data_a2 <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(fitted_values = fitted(model6)) %>%  # Add fitted values from model 6 for Donald Trump
  group_by(state) %>%
  mutate(
    weight = numeric_grade * (max_pollscore - pollscore) /
             sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)
  ) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_trump = sum(fitted_values * weight, na.rm = TRUE))

@@

# Filter out Alaska, Hawaii, and Puerto Rico from state_support_data
map_data_continental <- subset(us_states, !NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

@@

# Define the colors for each category
color_scale <- c("Kamala Harris" = "steelblue", "Donald Trump" = "indianred", "No Polling Data" = "grey")

# Plotting the map with ggplot2
ggplot(map_data_continental) +
  geom_sf(aes(fill = winner), color = "white", size = 0.2) +
  geom_text(aes(label = STUSPS, geometry = geometry), 
            stat = "sf_coordinates", color = "black", size = 3) +  # Adjust text size as needed
  scale_fill_manual(
    values = color_scale, 
    na.value = "grey",
    name = "2024 Election Winner",
    labels = c("Kamala Harris", "Donald Trump", "No Polling Data")
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Adjust the aspect ratio to make the map wider
  )

@@

Swing states, like Pennsylvania, Michigan, and Wisconsin, are predominantly leaning toward Harris, indicating a possible advantage for her in these key battlegrounds, although other typical swing states like Arizona and Georgia lean toward Trump. Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections. This visualization highlights the geographical and political divides across the U.S., as well as the critical role swing states play in determining the election outcome.

@@

# Define battleground states based on provided list
battleground_states <- c("Pennsylvania", "Michigan", "North Carolina", "Wisconsin", "Nevada", "Arizona", "Georgia")

@@

# Display the table using kable
knitr::kable(battleground_table)

USAFacts [@usafacts], FiveThirtyEight [@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election. Swing states are crucial for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate. @tbl-battleground-states summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin. The data shows close margins in these states, with some like Arizona and Georgia leaning towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins. The support margin column highlights the competitiveness of these states, with all margins below 2.2%.

@@

Our state-level analysis highlights close competition observed in key battleground states, emphasizing their pivotal role in the 2024 election. Our model reveals that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these critical electoral votes. The predictions show that even minor shifts in support within these swing states could decisively impact the overall election outcome. This analysis shows the importance of regional polling in capturing nuanced voter sentiment and the heightened influence that battleground states hold in shaping the final results.

However, there are limitations to our approach. Our model assumes linear relationships between predictors and candidate support, which might not capture more complex or interaction-based trends. Additionally, while we weighted polls based on their numeric grades, these scores may not fully reflect each pollster’s accuracy, leaving room for improvement. Future research could explore non-linear methods, such as machine learning, to capture potential interactions between variables. Refining the weighting mechanism by considering pollster performance history or state-specific polling nuances could further enhance predictive accuracy.

@@

The Emerson survey is designed with simplicity and efficiency. One of its strengths is its straightforward format, using clear and unambiguous questions. This simplicity makes the survey easy for respondents to understand. Another strength is its topical relevance. By focusing on vote preferences and demographic splits, the survey provides timely insights, especially valuable during election cycles. However, the survey has some limitations. One is its lack of nuance; while it captures basic voter preferences, it doesn’t explore the motivations or factors driving those choices in depth. Additionally, the various survey modes can affect respondent engagement with the questions, which may also influence the accuracy and depth of their answers.

@@

Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach. Its weighting techniques contribute to reliable and accurate results despite non-responses. While the use of multiple survey modes reduces coverage bias, other possible biases—such as lack of nuance and combined measurement errors must be considered during the analysis.

@@

We would oversample key swing states like Pennsylvania, Arizona, and Georgia [@swing_state]. Key swing states are states where there is no overwhelming support for a particular candidate or political party [@barker]. The outcome in these states often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election: An Analysis of Battleground States"
subtitle: "Trump Favored Nationally, While Harris Leads in 4 of 7 Battleground States"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."
date: today
date-format: long
abstract: >
  This study employs a poll-of-polls approach to predict support for Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election, with a focus on key battleground states to forecast the likely winner of the electoral college. By aggregating multiple polls and applying a weighted linear regression model with predictors like pollster reliability, sample size, state, and recency, we estimate a higher national support for Trump relative to Harris. Our analysis also shows a close competition in battleground states, with Trump holding a narrow lead in Arizona, Georgia, and North Carolina, while Harris leads in Michigan, Nevada, Pennsylvania, and Wisconsin. These findings underscore the value of aggregating polls over relying on individual surveys, offering a robust forecast of electoral outcomes.
format: 
  pdf:
    number-sections: true
    header-includes:
      - \usepackage{pdfpages}
      - '\KOMAoption{captions}{tableheading}'
      - \usepackage[margin=1in]{geometry}

@@

This paper contributes to the field of election forecasting by emphasizing the importance of poll quality and recency in prediction models. By focusing on battleground states where the support margins are particularly close, our approach highlights the areas where campaign strategies and voter turnout efforts may have the greatest impact. This analyasis is useful for political strategists, media analysts, and the general public by offering a view of the electoral landscape.

@@

## Variables of Interest

@@

```{r, fig.width = 12, fig.height = 6}
#| warning: false
#| message: false
#| echo: false
#| label: fig-ster
#| fig-cap: 'Support distribution for Kamala Harris and Donald Trump by major pollsters, highlighting variability in reported support across organizations and the value of aggregating multiple polls.'

@@

@fig-pie illustrates the distribution of polls between candidates and poll types. The left chart shows that the majority of polls are conducted at the state level, with a smaller portion being national. The right chart shows that polling is almost evenly split between Kamala Harris and Donald Trump. This distribution underscores the model’s balanced approach to capturing state-level differences as well as broader national trends, providing a view of candidate support across different contexts

@@

# Filter out Alaska, Hawaii, and Puerto Rico using subset
map_data_continental <-
  subset(map_data,!NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

ggplot(map_data_continental) +
  geom_sf(aes(fill = kamala_proportion * 100),
          color = "white",
          size = 0.2) +
  geom_text(
    aes(label = STUSPS, geometry = geometry),
    stat = "sf_coordinates",
    color = "black",
    size = 3
  ) +  # Adjust color and size as needed
  scale_fill_distiller(
    palette = "RdBu",
    direction = 1,
    na.value = "grey",
    name = "Harris's Mean Support in Polls (%)"
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Make the plot wider
  )
```

@@

In our approach, we will employ linear modeling of voter support percentage (pct) on pollster and other independent variables such as sample size, poll recency, and poll scope (state vs. national). This will allow us to smooth out the inherent noise, biases, and variability across different pollsters. Once we obtain the predicted values from our model, we will weight these predictions based on the numeric grade (quality score) of each pollster to calculate an overall national estimate of the outcome. Additionally, we will separately compute estimates for key battleground states, where voter behavior can be more volatile and pivotal in deciding the final outcome of the election. This approach helps us capture both national trends and state-level dynamics.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-causal-model
#| fig-cap: "Factors influencing candidate support, with poll attributes (Pollster, Sample Size, State, Recency) as direct contributors and unobserved confounders representing potential biases in polling results."

# Load and display the saved PNG
knitr::include_graphics(here("other", "graphs", "polls_of_polls_dag.png"))

@@

In our model, we aim to smooth out discrepancies and biases across various polling organizations using a polls-of-polls approach. Given the potential for individual pollsters to introduce systematic differences—due to variations in sampling methods, question phrasing, and historical leanings—our model includes a pollster variable to adjust for these organization-specific biases. This allows us to capture an aggregated view of public support that is less susceptible to the idiosyncrasies of any single poll. Furthermore, we incorporate sample size as a predictor, as polls with larger samples tend to yield more stable and reliable results, reducing random fluctuations caused by smaller samples. The state variable accounts for regional political differences, ensuring that the model captures varying levels of support across geographic and demographic groups, which is important for understanding the political landscape. Additionally, recency is included to prioritize more recent polls, as public opinion can shift in response to political events, and recent data is generally more reflective of current sentiment. By integrating these factors, our model seeks to produce a more stable measure of candidate support, reducing noise from individual poll discrepancies and focusing on a balanced, up-to-date aggregation of polling data.

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: tbl-summary
#| tbl-cap: "Model performance summary showing improved fit and accuracy as State and Recency are added, with R² increasing from 0.375 in Model 1 to 0.774 in Model 3, and RMSE decreasing from 3.053 to 1.836."

# Load your models
model1 <- readRDS(here("models/model1.rds"))
model2 <- readRDS(here("models/model2.rds"))
model3 <- readRDS(here("models/model3.rds"))

# Extract summary statistics and add variables included in each model
model_summary <- tibble::tibble(
  Model = c("Model 1", "Model 2", "Model 3"),
  Variables = c(
    "Pollster, Sample Size",
    "Pollster, Sample Size, State",
    "Pollster, Sample Size, State, Recency"
  ),
  `R²` = c(
    summary(model1)$r.squared,
    summary(model2)$r.squared,
    summary(model3)$r.squared
  ),
  `Adjusted R²` = c(
    summary(model1)$adj.r.squared,
    summary(model2)$adj.r.squared,
    summary(model3)$adj.r.squared
  ),
  `AIC` = c(AIC(model1), AIC(model2), AIC(model3)),
  `BIC` = c(BIC(model1), BIC(model2), BIC(model3)),
  `RMSE` = c(sqrt(mean(
    residuals(model1) ^ 2
  )), sqrt(mean(
    residuals(model2) ^ 2
  )), sqrt(mean(
    residuals(model3) ^ 2
  )))
)

@@

Aggregating support across all polls and applying our weighted approach, we estimate the overall support for each candidate. The weights, calculated based on both the poll's numeric grade and poll score, ensure that higher-quality polls with less bias contribute more to our estimates. Based on this approach, Kamala Harris has an estimated overall weighted support of around 47.77%, while Donald Trump stands at around 46.31%. This aggregation takes into account the varied methodologies and sampling qualities of different polling organizations.

@@

# Step 1: Subset data for Kamala Harris, calculate weights, join with fitted values from model3, and calculate weighted_pct
kamala_data <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
      sum(numeric_grade * (
        max(analysis_data$pollscore, na.rm = TRUE) - pollscore
      ), na.rm = TRUE),
    fitted_values = fitted(model3),
    # Replace this with actual fitted values for Kamala
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Kamala
  )

@@

# Step 2: Subset data for Donald Trump, calculate weights, join with fitted values from model6, and calculate weighted_pct
trump_data <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(
    weight = numeric_grade * (max(analysis_data$pollscore, na.rm = TRUE) - pollscore) /
      sum(numeric_grade * (
        max(analysis_data$pollscore, na.rm = TRUE) - pollscore
      ), na.rm = TRUE),
    fitted_values = fitted(model6),
    # Replace this with actual fitted values for Trump
    weighted_pct = fitted_values * weight  # Calculate weighted percentage for Trump
  )

# Calculate the overall weighted prediction for Donald Trump
trump_weighted_support <- sum(trump_data$weighted_pct, na.rm = TRUE)
```

@@

For each state, we compare the predicted support percentages and determine the projected winner based on which candidate has higher weighted support. We also calculate the margin of difference between the candidates, which shows how close each state’s race is.

@@

# Step 1: Subset data for Kamala Harris, join fitted values, then calculate weights within each state
kamala_data_a2 <- analysis_data %>%
  filter(candidate_name == "Kamala Harris") %>%
  mutate(fitted_values = fitted(model3)) %>%  # Add fitted values from model 3 for Kamala Harris
  group_by(state) %>%
  mutate(weight = numeric_grade * (max_pollscore - pollscore) /
           sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_kamala = sum(fitted_values * weight, na.rm = TRUE))

# Step 2: Repeat for Donald Trump with model6
trump_data_a2 <- analysis_data %>%
  filter(candidate_name == "Donald Trump") %>%
  mutate(fitted_values = fitted(model6)) %>%  # Add fitted values from model 6 for Donald Trump
  group_by(state) %>%
  mutate(weight = numeric_grade * (max_pollscore - pollscore) /
           sum(numeric_grade * (max_pollscore - pollscore), na.rm = TRUE)) %>%
  # Calculate the overall weighted prediction for each state
  summarize(support_trump = sum(fitted_values * weight, na.rm = TRUE))

@@

# Filter out Alaska, Hawaii, and Puerto Rico from state_support_data
map_data_continental <-
  subset(us_states,!NAME %in% c("Alaska", "Hawaii", "Puerto Rico"))

@@

# Define the colors for each category
color_scale <-
  c(
    "Kamala Harris" = "steelblue",
    "Donald Trump" = "indianred",
    "No Polling Data" = "grey"
  )

# Plotting the map with ggplot2
ggplot(map_data_continental) +
  geom_sf(aes(fill = winner), color = "white", size = 0.2) +
  geom_text(
    aes(label = STUSPS, geometry = geometry),
    stat = "sf_coordinates",
    color = "black",
    size = 3
  ) +  # Adjust text size as needed
  scale_fill_manual(
    values = color_scale,
    na.value = "grey",
    name = "2024 Election Winner",
    labels = c("Kamala Harris", "Donald Trump", "No Polling Data")
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    aspect.ratio = 0.5  # Adjust the aspect ratio to make the map wider
  ) 
```

@@

Swing states, like Pennsylvania, Michigan, and Wisconsin, are predominantly leaning toward Harris, indicating a possible advantage for her in these key battlegrounds, although other typical swing states like Arizona and Georgia lean toward Trump. Gray-shaded states lack sufficient polling data to make a prediction, reflecting areas of data scarcity in the model's projections. This visualization highlights the geographical and political divides across the U.S., as well as the importance of swing states play in election outcome.

@@

# Define battleground states based on provided list
battleground_states <-
  c(
    "Pennsylvania",
    "Michigan",
    "North Carolina",
    "Wisconsin",
    "Nevada",
    "Arizona",
    "Georgia"
  )

@@

# Display the table using kable
knitr::kable(battleground_table)
```

USAFacts [@usafacts], FiveThirtyEight [@FiveThirtyEight] and many other polling websites are calling Arizona, Georgia, Michigan, Pennsylvania, Wisconsin, North Carolina, Nevada the 'swing states' of the 2024 presidential election. Swing states are important for election predictions because these states, with their historically close voting margins, often determine the overall outcome by tipping the electoral balance toward one candidate. @tbl-battleground-states summarizes the predicted support percentages for Kamala Harris and Donald Trump in key battleground states, along with the predicted winner and the support margin. The data shows close margins in these states, with some like Arizona and Georgia leaning towards Trump, while others such as Michigan and Pennsylvania favor Harris by narrow margins. The support margin column highlights the competitiveness of these states, with all margins below 2.2%.

@@

Our state-level analysis highlights close competition observed in key battleground states, emphasizing their pivotal role in the 2024 election. Our model shows that in states like Pennsylvania, Michigan, and Wisconsin, support levels for Harris and Trump are nearly tied, reflecting the intense contest for these electoral votes. The predictions show that even minor shifts in support within these swing states could decisively impact the overall election outcome. This analysis shows the importance of regional polling in capturing voter sentiment and the heightened influence that battleground states hold in shaping the final results.

However, there are limitations to our approach. Our model assumes linear relationships between predictors and candidate support, which might not capture more complex or interaction-based trends. Additionally, while we weighted polls based on their numeric grades, these scores may not fully reflect each pollster’s accuracy, leaving room for improvement. Future research could explore non-linear methods, such as machine learning, to capture potential interactions between variables. Refining the weighting mechanism by considering pollster performance history or state-specific polling differences could further enhance predictive accuracy.

@@

The Emerson survey is designed with simplicity and efficiency. One of its strengths is its straightforward format, using clear and unambiguous questions. This simplicity makes the survey easy for respondents to understand. Another strength is its topical relevance. By focusing on vote preferences and demographic splits, the survey provides valuable information during election cycles. However, the survey has some limitations. One is the lack of deeper understanding; while it captures basic voter preferences, it doesn’t explore the motivations or factors driving those choices in depth. Additionally, the various survey modes can affect respondent engagement with the questions, which may also influence the accuracy and depth of their answers.

@@

Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach. Its weighting techniques contribute to reliable and accurate results despite non-responses. While the use of multiple survey modes reduces coverage bias, other possible biases—such as lack of deeper understanding and combined measurement errors must be considered during the analysis.

@@

We would oversample key swing states like Pennsylvania, Arizona, and Georgia [@swing_state]. Key swing states are states where there is no overwhelming support for a particular candidate or political party [@barker]. The outcome in these states often determine the overall result of the election, making them key in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which is important for predicting the Electoral College outcome.
